Automated Topic Naming

### 2012-01-20 ese/ese.tex
latest edits

We wanted to see whether developers in the PostgreSQL project\footnote{Since the MySQL and MaxDB data had poor records for developer ids, we focused on PostgreSQL.} worked on similar NFRs. For example, one developer might be a usability expert, while
another developer is focused on security. We assume that the NFR labels associated with a commit implicitly reflect work on those NFRs. Being able to report on where effort is being expended is useful for tracking productivity (among other things).
%Within this section we focused on PostgreSQL.
We explored this issue by extracting, for each topic in the PostgreSQL data-set, the developer names responsible for the commits used to generate that topic, and 
the NFRs that are associated with the topic. 
The analysis in this section used Neil's annotations.
This produced a map which associated developer name with NFR labels. The six NFR labels (\emph{maintainability, usability, efficiency, functionality, portability, reliability}) and the \emph{None} label describe a 7-dimensional space, into which we can position developers, based on the frequency with which each label occurs in the map.

@@

Based on Figure \ref{fig:authorsim} we wanted to see which clusters
\texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
evaluated the clusters proportionately, that is if we use distance
measures that normalize the data and ignore magnitude (e.g. turning an
author's NFR contributions into a histogram or unit-vector) such as
cosine distance, Euclidean distance of unit vectors, or Pearson
correlation. Euclidean distance uses both angle and magnitude of a
vector. For all three non-magnitude distances measures we tend to see that
the three largest contributors, \texttt{momjian}, \texttt{tgl}, and
\texttt{petere} inhabit the same clusters when we use 2-clusters or 6-clusters. This confirms the observations in
Figure \ref{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
proportionally similar.

@@

To evaluate our results empirically, we compared them to the annotations that would result if the seven labels
 were applied randomly. In the first simulation, as shown by Figure \ref{fig:irr},
we sampled with replacement from our own NFR label distributions to
produce random ratings that looked like our own. We then applied the
Kappa-statistic on this sampling and our labelling, and repeated
this 100,000 times. We then compared this distribution of simulated
IRR ratings against our our IRR scores. We found that
for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
IRR measures 96\% of the time. For \emph{portability} and \emph{efficiency}, our IRR
was greater 100\% of the time both for Neil's ratings, Abram's ratings
and the union of both Neil and Abram. 
We know this by evaluating our IRR ratings against the empirical
cumulative distribution function (ECDF) of the IRR ratings of the 100,000
simulation runs.
One explanation of the positive result for \emph{portability} is that such changes are often 
accompanied by clear indicator terms such as \emph{Windows, Linux, SunOS}, etc.

@@

PostgreSQL's size in terms of commits was equal to MySQL and MaxDB
combined. This imbalance in size, combined with the choice of 20
topics per month produced topics that represented too many issues or
threads within PostgreSQL's development. Choice of number of topics
should probably be tuned to the project.

@@

%XXX If we have more time try to elaborate more on this
One potential issue with some of our results is that we achieved very
low ROC and F-Measure scores.  Performance was not uniformly low,
as the more common classes often exhibited better performance than the
more rare classes.  Often this was amplified by class imbalance. In
future work we plan to investigate techniques such as sub-sampling and
boot-strapping in order to improve performance. Some of the ROC
performance could be due to a lack of coherence in tagging which was
shown by our low inter-rater reliability score. Inter-rater reliability is a threat, and we discussed it in Section \ref{sec:inter-rater}. 
% We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
% on the appropriate annotation for that topic. We verified IRR using
% manually annotated PostgreSQL topics, and achieved a weak result of $0.1$.
% Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
% usability or portability?). A short education session on what a given
% NFR from the ISO standard encompasses would reduce this error rate.
% Weak IRR results could also be responsible for some the low ROC
% performance, if each rater is picking up on different information it
% might make it more difficult to train a machine learner to do the
% same.
% Furthermore weak IRR results threatens the validity of the
% experiments, if raters cannot agree then consistently labelling by
% ISO-9126 related NFRs could be quite difficult.

>>>>

We wanted to see whether developers in the PostgreSQL project\footnote{Since the MySQL and MaxDB data had poor records for developer ids, we focused on PostgreSQL.} worked on similar NFRs. For example, one developer might be a usability expert, while
another developer is focused on security. We assume that the NFR labels associated with a commit implicitly reflect work on those NFRs. Being able to report on where effort is being expended is useful for tracking productivity (among other things).
%Within this section we focused on PostgreSQL.
We explored this issue by extracting, for each topic in the PostgreSQL data-set, the developer names responsible for the commits used to generate that topic, and 
the NFRs that are associated with the topic. 
The analysis in this section used Neil's annotations.
This produced a map which associated developer name with NFR
labels. The six NFR labels (\emph{maintainability, usability,
  efficiency, functionality, portability, reliability}) and the
\emph{None} label describe a 7-dimensional space (essentially a distribution), into which we can position developers, based on the frequency with which each label occurs in the map.

@@

Based on Figure \ref{fig:authorsim} we wanted to see which clusters
\texttt{momjian}, \texttt{tgl}, and \texttt{petere} would be in if we
evaluated the clusters proportionately, that is if we use distance
measures that normalize the data and ignore magnitude (e.g. turning an
author's NFR contributions into a histogram or unit-vector) such as
cosine distance, Euclidean distance of unit vectors, or Pearson
correlation distance ($1 - r$). Whereas, Euclidean distance uses both angle and magnitude of a
vector. For all three non-magnitude distances measures we tend to see that
the three largest contributors, \texttt{momjian}, \texttt{tgl}, and
\texttt{petere} inhabit the same clusters when we use 2-clusters or 6-clusters. This confirms the observations in
Figure \ref{fig:authorsim}, which uses Pearson correlation, that the larger contributors tend to be
proportionally similar.

@@

To evaluate our results empirically, we compared them to the annotations that would result if the seven labels
 were applied randomly. In the first simulation, as shown by Figure \ref{fig:irr},
we sampled with replacement from our own NFR label distributions to
produce random ratings that looked like our own. We then applied the
Kappa-statistic on this sampling and our labelling, and repeated
this 100,000 times. We then compared this distribution of simulated
IRR ratings against our own IRR ratings. We found that
for 4 out of 7 NFR labels, the IRR values for our labels were greater than the random
IRR measures 96\% of the time. For \emph{portability} and \emph{efficiency}, our IRR
was greater 100\% of the time both for Neil's ratings, Abram's ratings
and the union of both Neil and Abram. 
We know this by evaluating our IRR ratings against the empirical
cumulative distribution function (ECDF) of the IRR ratings of the 100,000
simulation runs.
One explanation of the positive result for \emph{portability} is that such changes are often 
accompanied by clear indicator terms such as \emph{Windows, Linux, SunOS}, etc.

@@

PostgreSQL was larger than MySQL and MaxDB. This imbalance in size,
combined with the choice of 20 topics per month produced topics that
represented too many issues or threads within PostgreSQL's
development. Choice of number of topics should probably be tuned to
the project.

@@

%XXX If we have more time try to elaborate more on this
One potential issue with some of our results is that we often achieved very
low ROC and F-Measure scores.  Performance was not uniformly low,
as the more common classes often exhibited better performance than the
more rare classes.  Often this was amplified by class imbalance. In
our future work we plan to investigate techniques such as sub-sampling and
boot-strapping in order to improve performance. Some of the ROC
performance could be due to a lack of coherence in tagging which was
shown by our low inter-rater reliability score. Inter-rater reliability is a threat, and we discussed it in Section \ref{sec:inter-rater}. 
% We assessed inter-rater reliability (IRR), a measure of the correlation between the manual annotations the first two authors created. Perfect IRR (1.0) indicates complete agreement 
% on the appropriate annotation for that topic. We verified IRR using
% manually annotated PostgreSQL topics, and achieved a weak result of $0.1$.
% Some of the reasons for this are a disagreement over the appropriate classification for documentation topics (should this be considered
% usability or portability?). A short education session on what a given
% NFR from the ISO standard encompasses would reduce this error rate.
% Weak IRR results could also be responsible for some the low ROC
% performance, if each rater is picking up on different information it
% might make it more difficult to train a machine learner to do the
% same.
% Furthermore weak IRR results threatens the validity of the
% experiments, if raters cannot agree then consistently labelling by
% ISO-9126 related NFRs could be quite difficult.
